\textbf{Species and genus plant classification with NEON hyperspectral remote sensing data}

### 2020-11-10 analysis/paper/paper.Rmd
intro text

NEON provides a unique opportunity to test, develop, and tune tree classification methods over a continental scale and at multiple taxonomic resolutions. 
We have extracted all NEON AOP hyperspectral data for every mapped stem in NEON plots where field data and airborne remote sensing data were collected during the same year.
We used a machine learning approach to evaluate plant identification potential using the hyperspectral data at multiple taxonomic resolutions: species and genus levels. 
Our study contributes to the active field of tree species classification research using freely available NEON data and open, reproducible code [@marconi2019data; @fricker2019convolutional; @scholl2020integrating]

@@

Each series of wavelengths increments by 5nm from each starting wavelength to yield 426 spectral bands. 
These starting wavelengths vary depending on which AOP payload was used to collect the hyperspectral data.
The wavelength values also change slightly each year when the NEON AOP re-calibrates their payload sensors [@kampe2010advances]. 
Most of them started at 384nm, so we linearly interpolated each spectrum to have the same 426 bands starting at 384nm and ending at 2512nm with 5nm increments.
Spectra from the Pu'u Maka'ala Natural Area Reserve (PUUM) NEON site in Hawaii had a starting wavelength of 347nm.
This is because NEON sub-contracted the PUUM flights to the Global Airborne Observatory, which has a spectrometer with a different wavelength range compared to the NIS payloads *([how to cite?](https://www.neonscience.org/observatory/observatory-blog/remote-sensing-partnership-made-paradise))*. 
We filtered out any spectra with a starting wavelength of 347nm, since the interpolation would require too much unavailable data outside the available range.
We flagged "bad bands" from 1340-1445nm, 1790-1995nm, and greater than 2400nm so they could be excluded from later analysis.
These wavelength ranges feature noisy reflectance data due to high water absorption [@bajorski2011].

Next, we removed samples belonging to species and genus classes with few samples. 
We figured that a small sample size would lead to a poor representation in model training. 
From the initial collection of 151 unique scientific names identified at the genus or species level, 50 of them had at least 20 samples and were kept for the classification analysis (Figure \@ref(fig:fig-histSpecies)). 
We randomly assigned each spectrum to the training or validation sets, using an 80/20 split. All 50 scientific name labels were represented in each set.

@@

We trained a random forest (RF) classifier to predict species based on NEON hyperspectral reflectance data as the descriptive features.
RF is a machine learning approach that utilizes an ensemble of decision trees and iterative subsets of the training data.
We performed hyperparameter optimization of *ntree*, the number of decision trees in the forest, and *mtry*, the number of variables randomly sampled at each node in the tree.
We used the optimal values of 27 for *mtry* and 1000 for *ntree* in our final RF model.
There were 49 species classes and 27 genus classes for each classification. 
We computed overall accuracy, confusion matrices, and variable importance plots.

>>>>

NEON provides a unique opportunity to test, develop, and tune tree classification methods over a continental scale and at multiple taxonomic resolutions. 
We have extracted all NEON AOP hyperspectral data for every mapped stem in NEON plots where field data and airborne remote sensing data were collected during the same year.
We used a machine learning approach to evaluate plant identification potential using the hyperspectral data at the species and genus taxonomic levels. 
Our study contributes to the active field of tree species classification research using freely available NEON data and open, reproducible code [@marconi2019data; @fricker2019convolutional; @scholl2020integrating]

@@

Each series of wavelengths increments by 5nm from each starting wavelength to yield 426 spectral bands. 
These starting wavelengths vary depending on which AOP payload was used to collect the hyperspectral data.
The wavelength values also change slightly each year when the NEON AOP re-calibrates their payload sensors [@kampe2010advances]. 
Most of them started at 384nm, so we linearly interpolated each spectrum to have the same 426 bands starting at 384nm and ending at 2512nm with 5nm increments.
Spectra from the Pu'u Maka'ala Natural Area Reserve (PUUM) NEON site in Hawaii had a starting wavelength of 347nm.
This is because NEON sub-contracted the PUUM flights to the Global Airborne Observatory, which has a spectrometer with a different wavelength range compared to the NIS payloads [@kampe2010advances]. 
We filtered out any spectra with a starting wavelength of 347nm, since the interpolation would require too much unavailable data outside the available range.
We flagged "bad bands" from 1340-1445nm, 1790-1995nm, and greater than 2400nm so they could be excluded from later analysis.
These wavelength ranges feature noisy reflectance data due to high water absorption [@bajorski2011].

Next, we removed samples belonging to species and genus classes with few samples, since extreme class imbalance is a known problem that can lead to biased classification results [@bader2018biased]. 
From the initial collection of 151 unique scientific names identified at the genus or species level, 50 of them had at least 20 samples and were kept for the classification analysis (Figure \@ref(fig:fig-histSpecies)). 
We randomly assigned each spectrum to the training or validation sets, using an 80/20 split. All 50 scientific name labels were represented in each set.

@@

We trained a random forest (RF) classifier to predict species based on NEON hyperspectral reflectance data as the descriptive features.
RF is a machine learning approach that utilizes an ensemble of decision trees and iterative subsets of the training data [@breiman2001random].
We performed hyperparameter optimization of *ntree*, the number of decision trees in the forest, and *mtry*, the number of variables randomly sampled at each node in the tree.
We used the optimal values of 27 for *mtry* and 1000 for *ntree* in our final RF model.
There were 49 species classes and 27 genus classes for each classification. 
We computed overall accuracy, confusion matrices, and variable importance plots.
